%%---------------------------------------------------------------------------%%
%% draco-autodoc.tex
%% John McGhee
%% $Id$
%%---------------------------------------------------------------------------%%
\documentclass[11pt]{nmemo}
\usepackage[centertags]{amsmath}
\usepackage{amssymb,amsthm,graphicx}
\usepackage[mathcal]{euscript}
\usepackage{tmadd,tmath}
\usepackage{c++}
\usepackage{cite}

%%---------------------------------------------------------------------------%%
%% DEFINE SPECIFIC ENVIRONMENTS HERE
%%---------------------------------------------------------------------------%%
%\newcommand{\elfit}{\ensuremath{\operatorname{Im}(-1/\epsilon(\vq,\omega)}}
%\msection{}-->section commands
%\tradem{}  -->add TM subscript to entry
%\ucatm{}   -->add trademark footnote about entry

\newcommand{\draco}{{\normalfont\sffamily Draco}}
\newcommand{\solon}{{\normalfont\sffamily Solon}}
\newcommand{\milagro}{{\normalfont\sffamily Milagro}}
\newcommand{\xemacs}{{\normalfont\bfseries XEmacs}}
\newcommand{\gnats}{{\normalfont\bfseries GNATS}}
\newcommand{\doxy}{{\normalfont\bfseries Doxygen}}
\newcommand{\imc}{{\normalfont\sffamily imc}}
\newcommand{\dsxx}{{\normalfont\sffamily ds++}}
\newcommand{\cfour}{{\normalfont\sffamily c4}}
\newcommand{\rng}{{\normalfont\sffamily rng}}
\newcommand{\timestep}{{\normalfont\sffamily timestep}}

%%---------------------------------------------------------------------------%%
%% BEGIN DOCUMENT
%%---------------------------------------------------------------------------%%

\begin{document}

%%---------------------------------------------------------------------------%%
%% OPTIONS FOR NOTE
%%---------------------------------------------------------------------------%%

\toms{Distribution}
%\toms{Joe Sixpak/X--TM, MS D409}
\refno{XTM:JMM-99-57 (U)}
\subject{\doxy\ \C++ Automatic Documentation Utility}

%-------NO CHANGES
\divisionname{Applied Theoretical \& Computational Physics Div.}
\groupname{X--TM:Transport Methods Group}
\fromms{John M. McGhee/X--TM, MS D409\\
  Tom M. Evans/X--TM, MS D409}
\phone{(505)667--9552}
\originator{jmm}
\typist{jmm}
\date{\today}
%-------NO CHANGES

%-------OPTIONS
%\reference{NPB Star Reimbursable Project}
%\thru{P. D. Soran, XTM, MS B226}
%\enc{list}      
%\attachments{list}
%\cy{list}
%\encas
%\attachmentas
%\attachmentsas 
%-------OPTIONS

%%---------------------------------------------------------------------------%%
%% DISTRIBUTION LIST
%%---------------------------------------------------------------------------%%

\distribution {
  J.E. Morel, XTM MS D409\\
  J.M. McGhee, XTM MS D409\\
  H.G. Hughes, XTM MS D409\\
  T.M. Evans, XTM MS D409\\
  M.G. Gray, XTM MS D409\\
  S.D. Pautz, XTM MS D409\\
  R.M. Roberts, XTM MS D409\\
  T.J. Urbatsch, XTM MS D409\\
  T.A. Wareing, XTM MS D409\\
  J.S. Warsa, XTM MS D409\\
  C.J. Gesh, XTM MS D409\\
  W.D. Hawkins, XTM MS D409\\
  R.B. Lowrie, XHM MS D413 
  }

%%---------------------------------------------------------------------------%%
%% BEGIN NOTE
%%---------------------------------------------------------------------------%%

\opening

The GNU Public License \C++ automatic code documentation system \doxy\ 
has been installed for use by the \draco\ team. This memo describes
how to set up and use \doxy\ .  Also, some suggestions are made as to
minimal documentation standards.

%%---------------------------------------------------------------------------%%

\section{Introduction}

\doxy\ is an automatic \C++ code documentation system available under
the GNU Public License.  A user's manual and other information is
available on-line at:
\begin{verbatim} 
     http://www.stack.nl/~dimitri/doxygen/ 
\end{verbatim}

Good documentation is critical to any code development effort.  The
advantages of automatic documentation systems are many.  For example,
the documentation and the code are much more likely to be conformal;
the documentation is conveniently located with the code, code
developers are encouraged to include liberal explanations in their
comments, and updates and corrections are easily handled.  \draco\ has
been considering adopting such a system for some time.


Recently, Tom Evans has evaluated several automatic documentation
systems including DOC++, PERCEPS, KDOC, and \doxy\ .  For \draco\ 
automatic documentation purposes, \doxy\ is deemed to be clearly the
package of choice. There are several reasons for this. Among these are
ease of use and installation, a plethora of clever and useful
features, good documentation (the package documented itself!), active
development and support from a responsive author, and minimal overhead
requirements on the code developer.


This evaluation is based on direct experience using \doxy\ to document
two existing \draco\ packages.  Tom Evans has documented the rng package and
John McGhee has documented the timestep package.  

%%---------------------------------------------------------------------------%%

\section{How to use \doxy\ }

\subsection{Overview}

\doxy\ can be used on any C++ source file
without any modifications to the source at all.  
\doxy\ groks \C++ and extracts considerable information from the *.hh
file without any active intervention on the part of the developer.
In addition to this information, the developer may also flag certain
comment blocks for inclusion in the documentation, if desired.
The only additional burden this places on the code author is to 
remember and occasionally use a dozen or so
simple comments. Both html (with links)
and \LaTeX\ output is automatically produced directly from the source code.
Example code can be included verbatim. If desired, both  html tags and \LaTeX\
formulas can be included directly in the source code comments. 

\subsection{Special Comment Blocks}
\label{sec:spcom}

If the developer wishes to provide extra commentary, 
a few simple special comment characters are available to extract additional 
documentation directly from the source code comment blocks.
The fundamental key that \doxy\ uses to flag source code text for
extraction is the C-style comment character followed by an exclamation
point. For example:
\begin{verbatim}
     /*!
      *
      * This text (without the asterisks) will appear in the documentation!
      *
      */
\end{verbatim}
This is all that is required from the code developer to get everything
in the comment block into the documentation. \doxy\ automatically associates
the special comment block with the class/member that immediately follows the 
comment block in the source. Alternatively, you can put the comment block
anywhere in any file and mark it with a special keyword to associate it
with the appropriate class/member. Blank lines are recognized as paragraph
breaks. Code comments in other legal \C++ comment formats are ignored.

\subsection{A Few Major Keywords}

Inside a \doxy\ comment block, there are currently approximately 60
special keywords that \doxy\ recognizes. If desired, these keywords
can be used to control and produce additional features in the
documentation.  The interested reader is referred to the \doxy\ manual
for details on these keywords.  We will briefly describe a few of the major
keywords that we have found to be most useful.
\begin{itemize}
\item \verb+\file+ The file name.
\item \verb+\author+ The author.
\item \verb+\date+ The date.
\item \verb+\brief+ A pithy ``one line'' description.
\item \verb+\param+ An input parameter.
\item \verb+\return+ The return value.
\item \verb+\sa+ ``See Also'', a list of related items.
\item \verb+\example+ {\em filename} Includes {\em filename} as a code
  example.
\item \verb+\include+ {\em filename} Includes {\em filename} verbatim at the
  point of reference.
\item \verb+\page+ {\em reference title} Places the comment block in which
  it appears in the documentation as a separate page with title {\em
    title}. You can refer to and create a link to the page using {\em
    reference} as an argument to the \doxy\ ``\verb+\ref+'' keyword.
\end{itemize}

\subsection{Other Fancy Stuff}

If found within a special comment block 
as defined in section \S~\ref{sec:spcom} above, \doxy\ 
recognizes a multitude of html commands. These
commands are used directly in the html document, and are converted appropriately for
the \LaTeX\ output. 
For example, if the phrase: 
\begin{verbatim}
<H1> My Header <\H1>
\end{verbatim} appears in the special comment
block \doxy\ will format the phrase ``My Header'' as the start of an
un-numbered section in the documentation. 
Also, \LaTeX\ formulas can be used inside a special
comment block if they are delimited using the syntax ``\verb+\f$ ... \f$+''.
For example, the phrase 
\begin{verbatim}
\f$ \sqrt{x_2-x_1)^2+(y_2-y_1)^2} \f$ 
\end{verbatim}  will appear in the
documentation as $\sqrt{x_2-x_1)^2+(y_2-y_1)^2}$.
These two
features make it possible to create a fairly sophisticated document, if desired.
See the manual for more details.


%%---------------------------------------------------------------------------%%

\section{\draco\ Installation Strategy}
\label{sec:draco}

\subsection{File Locations}
We will create a \texttt{draco/autodoc} directory. Under this directory
will be a \doxy\ configuration file (\texttt{draco.config)}, html header 
and footer files,
and a tagfile (\texttt{draco.tag}). Also in this directory will be found
a \texttt{draco/autodoc/html} directory to store \doxy\ html file
output and a \texttt{draco/autodoc/latex}
directory to hold \doxy\ \LaTeX\ output. 

\subsection{Usage}
To run \doxy\ simply cd into the \texttt{draco/autodoc} directory and type:
\begin{verbatim} 
     PATH/doxygen draco.config
\end{verbatim}
Currently ``PATH'' is \texttt{/scratch/tme/bin} and \doxy\
can only be executed from a SGI platform. These details may
change in the future. 
A html and \LaTeX\ directory will be produced with your documentation 
inside. 

\subsection{\doxy\ Configuration File}
\label{sec:config}

Under normal use the \doxy\ configuration file will not require
modification by package developers. However, if new packages or
new directories under existing packages are added,
the \doxy\ configuration file \texttt{draco/autodoc/config.draco}
may require modification.
The configuration options available to users are well
documented in the \doxy\ manual. Some of the configuration file 
variables which may be of interest are briefly discussed below:

\begin{description}
\item[\ttfamily PROJECT\_NUMBER] This tag gives the release number of
  the documentation.  It must be updated when the documentation is released.
  See  Ref~\cite{xtm:9936} for details on release numbers.
\item[\ttfamily INPUT] The user must define which files and/or
  directories should be  examined by \doxy\ for documentation.  
\item [\ttfamily FILE\_PATTERNS] One or more wildcard patterns to be
  used as filters for files found in the INPUT directories.
\item [\ttfamily EXCLUDE] Specifies files and/or directories to be
   excluded from the INPUT source files.
\item [\ttfamily EXCLUDE\_PATTERNS] One or more wildcard patterns to be
  excluded from the INPUT directories.
\item [\ttfamily EXAMPLE\_PATH] This tag lists one or more files and/or
  directories that contain example code.
\item [\ttfamily INCLUDE\_PATH] This tag lists one or more 
  directories that contain files to be included in the documentation.

\end{description}

\subsection{Release and Tagging Procedures}
Prior to releasing a \draco\ package, 
\doxy\ should be run, thus assuring that the documentation is in sync
with the source code. Alternatively, we could include documentation
production as part of the \draco\ build, but this would require
\doxy\ to be installed on all platforms on which we expect to build
\draco.  We feel that this too restrictive. Also, we feel that
the package documentation should be available to the user immediately
upon checkout.

\subsection{New Templates and Files}

Tom Evans is installing new *.cc and *.hh templates in the \draco\ 
templates directory to support \doxy\ requirements. These templates
will be available in the usual manner as described in Ref~\cite{xtm:9909}
There will also be
a \doxy\ configuration file, and standard
header and footer files for \doxy\ html output installed in the 
\texttt{draco/autodoc} directory.
 

\subsection{Routine Maintenance of \doxy\ Output}


 \doxy\ will need to
be run periodically to bring the documentation into sync with with the
source code. 
Each package developer will be responsible for running
 \doxy\ when their packages are updated.
Also the version number in the \texttt{draco/autodoc//config.draco}
file will need to be periodically updated to reflect the current version number.
The resulting output should be tagged and committed to
CVS using the same procedures as for the source code. Note that
\doxy\ may create gif images as part of the html generation
process. To prevent possible difficulties with CVS, these image files
 should be added to the repository using the CVS ``-kb'' option. 

\subsection{\draco\ WWW Site}

When checking out the \draco\ web site, CVS magic will be used to
check out the appropriate html source in
\draco. Stubs for links to on-line documentation have already been
installed on the web site.


%%---------------------------------------------------------------------------%%

\section{Documentation Standards}

We believe that each developer should be free to document source code
in the style and manner most agreeable to the individual. As a result,
only minimal documentation standards are required.  We propose the
following standards for consideration:

\begin{itemize}
\item Each file, namespace, class, and associated public members
  should have at least a brief description.
\item Each public member should have descriptions of its input
  parameters and return values.
\end{itemize}

We feel that these guidelines, together with the information that
\doxy\ will automatically extract, will produce documentation that is
adequate for our purposes. Of course, all developers are free and
encouraged to provide more than the absolute minimum.

%%---------------------------------------------------------------------------%%

\section{Using \doxy\ with Other Systems}

The \draco\ \doxy\ documentation system can be used as a model for
the setup of automatic documentation of other software systems. 
Alteration will be 
required to the \texttt{<system>/autodoc/<system>.config} files, and
to the \texttt{html} header and footer files. 


\subsection{Linking to Draco HTML Documentation from Other Systems}

It is possible to link to \draco\ \texttt{html} documentation from other
software systems documentation.
Tag files are the  optional \doxy\ feature used to refer to other software systems. 
For example, the \milagro\
package uses the \draco\ package.  The documentation for \milagro\
may therefore provide links to the documentation for \draco.  If such is 
desired, it is  accomplished
by the \doxy\ tagging machinery.  To use \doxy\ tags the following
config options are used:
\begin{description}

\item[\ttfamily TAGFILES] A list of all tag files needed to resolve
  links in the package.  For instance, the \milagro\ package
  \texttt{autodoc/milagro.config} file would contain the following entry:
\begin{verbatim}
     TAGFILES = ./html/draco.tag
\end{verbatim}
  This is  the tag file used by \milagro\ to resolve documented
  dependencies. It should be copied into the \milagro\ \texttt{autodoc/html}
  directory from the \draco\ autodoc directory.

\item[\ttfamily GENERATE\_TAGFILE] forces the generation of
  a tag file in the \texttt{autodoc} subdirectory with the name 
  \texttt{<system>.tag}.
For instance, the \milagro\ package
  \texttt{autodoc/milagro.config} file would contain the following entry:
\begin{verbatim}
     GENERATE_TAGFILES = ./milagro.tag
\end{verbatim}

\end{description}
When \doxy\ is run with these configuration options, link information
is generated automatically to the \draco\ dependencies.  Additionally, 
\milagro\ creates its own tag file that can be used by its client
systems, if any.

When the \texttt{TAGFILES} option is set, \doxy\ creates a customized
PERL script called \texttt{installdox} in the \texttt{html}
subdirectory.  This script is used to ``activate'' the links to real
locations.  In order to use this script, copies of each tag file must
exist in the \texttt{html} subdirectory.  The \texttt{installdox}
script is then executed to modify the links to their proper locations.
At this point an example is warranted.  Consider the \milagro\ 
documentation set.  \milagro\ depends on \draco.  Thus, the following
steps are required to activate the tags in the \milagro\ documentation
that point to the \draco\ documentation:
\begin{verbatim}
     cd milagro/autodoc/html
     cp PATH/draco.tag .
     ./installdox -l draco.tag@LOCATION
\end{verbatim}
where ``PATH'' is the path to the \texttt{draco.tag} file 
and ``LOCATION'' is the
location of the draco \texttt{html} directory.
The \texttt{installdox} script will modify all of the \milagro\ \texttt{.html}
files so that the links point to the correct locations.    The dates of the
copied tag files
can be used to determine the currency of the references to other packages.
Of course, the \texttt{.html} files must actually exist at the stated location
for the links created by \texttt{installdox} to actually work.

We emphasize that the use of tag files is optional. It is somewhat cumbersome,
but provides a useful enhancement of the \texttt{html} output. It has no
affect on the \LaTeX\ output. For any who find tag files confusing or
intimidating, we recommend that you document your system without them. They
can easily be added at a later date if desired.


\subsection{Currency Considerations}
Note that the documentation for 
any system ``A'' that uses system ``B's'' B.tag file will have to be
updated when B.tag is changed. This update consists of getting the updated
tag file and running \texttt{installdox} again. 
Ideally, there should be some way to detect that documentation
or tag files are out of date, and to automatically rebuild the
documentation if required.  For the present at least, no such automatic
feature is in place, and we elect to rely
on the diligence and dedication of the individual package developers to 
assure that the documentation is kept up to date. 


%%---------------------------------------------------------------------------%%

\section{Example}

Nothing is as instructive as a good example, so here is a copy of
one of the header files from the \timestep\ utility. Of course,
there is more than one way to accomplish this, but this is one way.
Additionally, both the \rng\ and \timestep\ packages have been
autodocumented.  Feel free to use these as examples. 

\begin{verbatim}
//----------------------------------*-C++-*----------------------------------//
/*! 
 * \file fixed_ts_advisor.hh
 * \brief Defines the fixed time-step advisor.
 *
 * \author <a href="http://www.lanl.gov/home/mcghee">
 *  John McGhee</a>
 *
 * \date Thu Apr  2 14:06:18 1998
 */
//---------------------------------------------------------------------------//

#ifndef __timestep_fixed_ts_advisor_hh__
#define __timestep_fixed_ts_advisor_hh__

#include "ts_advisor.hh"

namespace rtt_timestep {

//===========================================================================//
/*!
 * \brief Introduces a user defined fixed value 
 *        into the time-step calculation.
 *
 * \sa The ts_manager class provides a means to manage this advisor.
 * The \ref overview page gives a summary of the Draco time step control
 * utilities. 
 *
 * This is useful to set min and max timesteps, or to force a
 * timestep, etc. The recommendation for the new timestep is
 * simply the user input value. 
 */ 
//===========================================================================//
class fixed_ts_advisor : public ts_advisor {

  // DATA

  private:
    /*!
     * \brief Value used to oompute a fixed advisor recommended timestep.
     */
    double fixed_value; 
   
// CREATORS

  public:

    /*!
     * \brief Constructs a fixed time-step advisor.
     * \param name_ A unique name for the advisor.
     * \param usage_ How the advisor is to be used.
     * \param const_value_ The desired value for the timestep.
     * \param active_ Turns the advisor on/off.
     */
    fixed_ts_advisor( 
        const std::string &name_  = std::string("Unlabeled"),
        const usage_flag usage_ = max, 
        const double const_value_ = large(),
        const bool active_ = true);

    /*!
     * \brief Destroys a fixed time-step advisor.
     */
    ~fixed_ts_advisor();

// MANIPULATORS
    
    /*!
     * \brief Sets the fixed value.
     * \param value_ The fixed value.
     */
    void set_fixed_value(const double value_ = large())
    { 
        fixed_value = value_;
    }

// ACCESSORS

    /*!
     * \brief Returns the time-step recommended by a fixed  advisor.
     * \param tsm The time step manager in which the advisor resides.
     * \return The recommended timestep.
     */
    double get_dt_rec(const ts_manager &tsm) const;
    
    /*! 
      \brief Prints the state of a fixed advisor.
      \return Prints the internal state of the advisor to std out.
     */
    void print_state() const;
    
    /*! 
      \brief Fixed advisor invariant function.
      \return True if the invariant is satisfied.
     */
    bool invariant_satisfied() const;

};

} // end of rtt_timestep namespace

#endif                          // __timestep_fixed_ts_advisor_hh__

//---------------------------------------------------------------------------//
//                              end of fixed_ts_advisor.hh
//---------------------------------------------------------------------------//

\end{verbatim}

%%---------------------------------------------------------------------------%%

\section{Closing}

In closing, we would like to emphasize that \doxy\ is
an easy and painless way to document your code. We have both
found the process addictive, the more comments you put in, the more
you want to put in. \doxy\ can be used on any \C++ source file without
any modifications to the source at all. However, we would encourage
all \draco\ developers to take the small amount of time required to
retrofit their existing code to take advantage of the special features
that \doxy\ offers.  Also, we recommend that all new code developed in
\draco\ make use of \doxy\ capabilites.

\bibliographystyle{rnote}
\bibliography{../bib/draco}

\closing
\end{document}

%%---------------------------------------------------------------------------%%
%% end of draco-autodoc.tex
%%---------------------------------------------------------------------------%%
